---
title: "Data Science for Bionformatics - project 3"
output:
  html_document:
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load libraries

```{r, message = FALSE}
library('tidyverse')
library('ggplot2') 
library('ggthemes') 
library('scales')
library('grid') 
library('gridExtra')
library('corrplot')
library('ggraph') 
library('igraph') 
library('dplyr') 
library('nycflights13')
library('readr') 
library('tibble') 
library('tidyr')
```

# Import data

```{r, message=FALSE, warning=FALSE}
bipol<- read.table(file="bipol2.csv",TRUE,";")
```

# Explore data
```{r}
summary(bipol)
head(bipol)
```
#3 groups of patients: people with bipolar disorder, people with depression and the control group
#I analysed the 4 types of people( there are 2 types of people with BD -> BD I and BD II, the control group and people with depression (UD))
#So the analysis is about the right answers from patients with different types of bipolar disorder

# Extract variables
```{r}
type=bipol$Type
answer=bipol$Right_answers
```

# Compare mean of right answers (BD I, BD II, Control, UD)
```{r}
tapply(answer, type, mean)
```
#null hypotesis: means of all the groups are equal / alternative hypotesis: at least 1 mean is different
#for more than 2 groups we no longer can use t-test, that's why I used ANOVA for my 4 groups 
#4 types of patients: 2 types of people with bipolar disorder (BDI, BDII) and people with depression(UD) and the control group

# Visualization
```{r}
boxplot(answer~type, main="Types of people with bipolar disorder and their right answers in the test",xlab="Types of patients",col=rainbow(4))
```

# Anova
```{r}
results=aov(answer~type)
summary(results)
```
#F-value: 39.57, p-value = 2e-16 < 0.05
#the alternative hypotesis is accepted, because the p-value is less than 0.05

# Post-hoc test
#it gives us which groups are significantly different from each other (<0.05) with 95% confidence level
```{r}
TukeyHSD(results)
```
#BD I & BD II types are not significantly different, but all the other types are very divergent, with 95% confidence level.

# Equality of variance
```{r}
bartlett.test(answer~type)
```
#the p-value is 0.02435 < 0.05, the null hypothesis in Bartlett's test states that all of the variances are equal and if we get their p-value of less than 0.05, we are able to reject the null hypothesis, so at least 2 of the types differ from each other, which is shown in the boxplot

